---
title: "hw3"
author: "Jason Pekos"
date: "27/09/2021"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(grid)
library(gridExtra)
library(ggthemes)
```


```{r, echo = FALSE}
Diseases = c("H3N2","H1N1","B","Cold", "Total")

DeathsNY    = c(1492,  -88,  774,   1640, 3819)
CINY_left   = c(361 , -569,  -21,  -1815,   66)
CINY_right  = c(2624,  384,  1571,  5097, 7572)
CI_dev_NY = abs(DeathsNY - CINY_right)
cityNY = "New York Metropolitan Area"


DeathsII    = c(2126,   127,  549,  1646, 4447)
CIII_left   = c(1004,  -338, -173, -2504,   62)
CIII_right  = c(3249,   592, 1271,  5796, 8832)
CI_dev_II = abs(DeathsII - CIII_right)
cityII = "Illinois and Indiana"



dfNY = tibble(Diseases,
              DeathsNY,
              CINY_left,
              CINY_right,
              cityNY,
              CI_dev_NY)
          
dfII = tibble(Diseases,
              DeathsII,
              CIII_left,
              CIII_right,
              cityII,
              CI_dev_II)
```



## Plots

For the first plot, I'll just do a standard CI-type plot for each region separately. I think this keeps clutter to a minimum, and emphasizes the subcategories in the original table. There are some other options here that I decided against:

\begin{itemize}
  \item Bar plots
  \item Connecting the points with \verb|geom_line()|
\end{itemize}

I didn't like bar plots here because I feel that just using \verb|geom_point()| emphasizes that this is a point estimate from a regression, and I feel that the bars add unnecessary clutter to the error bars. Also the presence of negative point estimates here doesn't look great on a bar plot (in my opinion).

I decided not to use \verb|geom_line()| because I feel it helps with absolutely nothing, even though I saw it used online a lot for this type of task. If there was some change over time, I would probably consider a super light version of it (thin mid-opacity dashed lines), but that isn't a part of this table. 

I made a few other minor additions --- most importantly, more breaks to help with interpretability. \

Overall though, there's only a few pieces of information we actually care about encoding here:

\begin{itemize}
  \item Geographic area studied
  \item Point estimate of Deaths
  \item Disease name
  \item Error estimates
\end{itemize}

We deal with the first by simply making two plots. We encode the second with position on a common scale, and we add the error estimates --- which are in the same units --- to the same scale. The only thing I'm unsure about is (not) encoding disease name with colour. I didn't end up doing this because it's already 'encoded' in some sense by position on the x-axis. It's a very simple table, and there isn't an interpretability issue with a graph like this. Especially with only five categories, I feel this would add needless clutter to a nice and simple graph.

```{r, echo = 0}
title = "Estimated Deaths in"

HW3Plotter = function(df){
  ggplot(df, aes (x = Diseases, y = pull(df, 2))) + 
    geom_point() +
    geom_errorbar(aes(ymin = pull(df, 2) - pull(df, -1),
                      ymax = pull(df, 2) + pull(df, -1)),
                  width  = 0.4)  +
    theme_hc() +
    scale_y_continuous(n.breaks = 8, limits  = c(-3000,10000)) +

    ggtitle(paste(title, pull(df, -2))) +
    xlab("Cause of Death")    +
    ylab("Deaths with 95% CI")    
}

HW3Plotter(dfNY)
cat("\\newpage")
HW3Plotter(dfII)
```




## Alternate Graph:

Simply seperating out the categories seems a little like cheating, so an alternate graph where we plot the same disease for both cities on the same graph, but spliting out each disease --- we do this because the wide range of estimates across diseases and the tiny variation within diseases makes differentiation super tough. 

```{r, echo = 0}
####CREATE NEW DFS FOR EASY PLOTTING 
#Probably what I should have done from the start
dfNY %>% rename(Deaths = DeathsNY,CI_left = CINY_left, CI_right = CINY_right, City = cityNY, CI_Dev = CI_dev_NY) -> dfNY2
dfII %>% rename(Deaths = DeathsII,CI_left = CIII_left, CI_right = CIII_right, City = cityII, CI_Dev = CI_dev_II) -> dfII2

dfFULL = rbind(dfNY2, dfII2)
```

```{r, echo = 0}
plots  = list()
#searching for something totally seperate, once I commited to this path already, I found this question
#https://stackoverflow.com/questions/9315611/grid-of-multiple-ggplot2-plots-which-have-been-made-in-a-for-loop
#which made my life here a lot easier. credit to this person (Justin).

for (Sick in unique(pull(dfFULL,1))) {
  plots[[Sick]] = ggplot(data = filter(dfFULL,  Diseases == Sick), aes(x = 0, y = Deaths, color = City)) +
              geom_errorbar(aes(ymin = Deaths - CI_Dev, ymax = Deaths + CI_Dev ),
                              position=position_dodge(0.05),
                            width = 0.3,
                            show.legend = FALSE) +
              geom_point(position=position_dodge(0.05),
                         show.legend = FALSE) +
    theme(aspect.ratio=18/9) +
    xlim(-1,1) +
    labs(x = Sick) + 
    theme(
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.title.y=element_blank()
        )

  
}

#more help 
#https://stackoverflow.com/a/49767811
#from user MBORG 

#also 
#https://stackoverflow.com/a/25962717
#user BAPTISTE 

#common legend
#https://stackoverflow.com/a/13650878
#credit to ROLAND 

main = do.call(grid.arrange,c(plots, top = "Estimated Deaths by Disease and Region", ncol = 5))

```

EXPLAIN POINT OF GRAPHS WHAT AM I TRYING TO SHOW BEFORE FRIDAY NIGHT

ALSO COMMENT CODE

LOOK HERE ALSO:

https://stackoverflow.com/a/49834950

